
```{r}
library(tidyverse)
library(synExtra)
library(fst)
library(data.table)
library(powerjoin)

synapser::synLogin()

syn <- synDownloader("~/data", .cache = TRUE)
```

```{r}
kinomescan_pseudo_kds <- syn("syn51080578") %>%
    read_csv()

kinomescan_classes <- syn("syn51080579") %>%
    read_csv()

kinomescan_kinase_info <- syn("syn51286743") %>%
    read_csv()

single_dose <- syn("syn26486828") %>%
    read_csv()

compound_mapping <- syn("syn26260389") %>%
  read_fst(as.data.table = TRUE)

# qpcr_misses <- syn("syn33531640") %>%
#   readxl::read_excel() %>%
#   left_join(
#     compound_mapping[, .(lspci_id, external_id)] %>%
#       unique(),
#     by = c("Compound Name" = "external_id")
#   )

# lsp_biochem <- syn("syn25173511") %>%
#   read_fst()
lsp_targets <- syn("syn25173506") %>%
  read_fst()

lsp_biochem_single <- syn("syn25173513") %>%
  read_fst()
```

Aggregate single-dose data to get IC50s. Remove any ChEMBL Kd values that 
have relation that is not "=" unless the relation is ">" or ">=" and the value
is greater than 9000. This is to avoid underestimating IC50s for cases where
people just didn't assay higher concentrations.

```{r}
chembl_kd_raw <- fread("data/kds_raw.csv.gz")

lsp_biochem <- lsp_biochem_single %>%
  filter(
    value_relation == "=" |
      (value_relation %in% c(">=", ">") & value > 9000)
  ) %>%
  group_by(lspci_id, lspci_target_id) %>%
  summarize(
    value = quantile(value, 0.25, na.rm = TRUE, names = FALSE),
    .groups = "drop"
  )
```

```{r pseudo_ic50_plots}

# pseudo_kd_class <- kinomescan_pseudo_kds %>%
#   power_full_join(
#     kinomescan_classes %>%
#       select(lspci_id, `DiscoveRx Gene Symbol`, classification) %>%
#       mutate(
#         classification = factor(
#           classification,
#           levels = c("binding", "weakly-binding", "undetermined", "non-binding")
#         )
#       ),
#     by = c("lspci_id", "DiscoveRx Gene Symbol"),
#     check = check_specs(
#       duplicate_keys_left = "abort",
#       duplicate_keys_right = "abort",
#       unmatched_keys_left = "abort",
#       unmatched_keys_right = "abort"
#     )
#   ) %>%
  # power_left_join(
  #   kinomescan_kinase_info %>%
  #     select(`DiscoveRx Gene Symbol`, hgnc_symbol),
  #   by = "DiscoveRx Gene Symbol",
  #   check = check_specs(
  #     duplicate_keys_left = "ignore",
  #     duplicate_keys_right = "abort",
  #     unmatched_keys_left = "abort",
  #     unmatched_keys_right = "ignore"
  #   )
  # ) %>%
  # power_left_join(
  #   lsp_targets %>%
  #     filter(organism == "Homo sapiens") %>%
  #     select(lspci_target_id, symbol),
  #   by = c("hgnc_symbol" = "symbol"),
  #   check = check_specs(
  #     duplicate_keys_left = "ignore",
  #     duplicate_keys_right = "ignore",
  #     unmatched_keys_left = "warn",
  #     unmatched_keys_right = "ignore"
  #   )
  # )
# NA ok. Comes from P. falciparum genes

class_colors <- c(
    "binding" = "firebrick1",
    "weakly-binding" = "darkorange",
    "undetermined" = "gray",
    "non-binding" = "black",
    "discordant" = "blue"
)

pseudo_kd_comp <- kinomescan_pseudo_kds %>%
  mutate(
    across(classification, \(x) factor(x, levels = names(class_colors)))
  ) %>%
  filter(dataset == "original_repeat_replaced") %>%
  power_left_join(
    lsp_targets %>%
      filter(organism == "Homo sapiens") %>%
      select(lspci_target_id, symbol),
    by = c("hgnc_symbol" = "symbol"),
    check = check_specs(
      duplicate_keys_left = "ignore",
      duplicate_keys_right = "ignore",
      unmatched_keys_left = "warn",
      unmatched_keys_right = "ignore"
    )
  ) %>%
  filter(
    !exclude_target
  ) %>%
  # group_by(
  #   lspci_id, name, hmsl_id, `Entrez Gene Symbol`, lspci_target_id
  # ) %>%
  # # Select representative for each gene among all
  # # mutants etc by favoring concordant measurements and low pseudo IC50s
  # arrange(classification == "discordant", pseudo_kd_relation != "=", pseudo_kd) %>%
  # slice_head(n = 1) %>%
  # ungroup() %>%
  inner_join(
    lsp_biochem,
    by = c("lspci_id", "lspci_target_id")
  )


```

Check weird cases where we have pseudo Kd < 1000 uM but binding
is classified as non-binding. Probably results mostly of cases
where entire binding curve is between 50-35% control so the 35% threshold
is never triggered but the 50% threshold for Kd is.

```{r}
pseudo_kd_comp %>%
    filter(pseudo_kd < 1000, classification == "non-binding", pseudo_kd_concordant == "concordant")

```

Only 152 such cases. Not worth worrying about.

Make scatterplot of ChEMBL IC50 vs KINOMEscan pseudo-Kd
Also show distribution of pseudo-IC50 values for each binding class as
beeswarm plot.

```{r}


p_scatter <- pseudo_kd_comp %>%
  drop_na(pseudo_kd) %>%
  filter(
    # pseudo_ic50_relation == "=",
    pseudo_kd_concordant == "concordant",
    classification != "discordant"
  ) %>%
  ggplot(
    # aes(value, pseudo_ic50, color = classification, shape = pseudo_ic50_relation)
    aes(
      value, pseudo_kd,
      color = classification,
      text = paste(
        name, `DiscoveRx Gene Symbol`,
        "<br>ChEMBL", signif(value, 2), "nM Eurofins", signif(pseudo_kd, 2), "nM"
      )
    )
  ) +
  scale_color_manual(values = class_colors) +
  geom_point(alpha = 0.5, shape = 16, size = 1) +
  scale_x_log10(labels = scales::label_number(big.mark = "")) +
  scale_y_log10(labels = scales::label_number(big.mark = "")) +
  coord_equal() +
  labs(
    x = "IC50 (ChEMBL)",
    y = "Pseudo-Kd (KINOMEscan)"
  )

ggsave(
    "plots/chembl_ic50_vs_kinomescan_pseudo_kd.pdf", p_scatter,
    width = 8, height = 3
)

library(plotly)

p_scatter_plotly <- ggplotly(p_scatter, tooltip = "text") %>%
  toWebGL()

htmlwidgets::saveWidget(p_scatter_plotly, "plots/chembl_ic50_vs_kinomescan_pseudo_kd.html")

library(ggbeeswarm)
library(ggtext)
library(glue)

class_labels <- glue_data(
  distinct(pseudo_kd_comp, classification) %>%
    filter(classification != "discordant"),
  "<span style='font-weight: bold; color: {class_colors[classification]}'>{classification}</span>"
) %>%
  set_names(
    distinct(pseudo_kd_comp, classification) %>%
      filter(classification != "discordant") %>%
      pull(classification)
  )

p_beeswarm <- pseudo_kd_comp %>%
  filter(
    pseudo_kd_concordant == "concordant",
    # pseudo_ic50_relation == "=",
    classification != "discordant"
  ) %>%
  mutate(
    classification = factor(
      classification,
      levels = names(class_colors) %>%
        rev()
    )
  ) %>%
  ggplot(
    aes(value, classification, color = classification)
  ) +
  geom_quasirandom(groupOnX = FALSE, shape = 16, size = 1, alpha = 0.5) +
  geom_boxplot(fill = NA, outlier.shape = NA) +
  scale_color_manual(values = class_colors) +
  guides(color = "none") +
  scale_x_log10(labels = scales::label_number(big.mark = "")) +
  scale_y_discrete(labels = class_labels) +
  theme(axis.text.y = element_markdown()) +
  labs(x = "IC50 (ChEMBL)", y = NULL)

ggsave(
    "plots/chembl_ic50_by_kinomescan_class_beeswarm.pdf", p_beeswarm,
    width = 7, height = 3
)

library(patchwork)
# p_combined <- wrap_plots(
#     p_scatter +
#         labs(x = NULL) +
#         theme(
#             axis.text.x = element_blank(),
#             plot.margin = unit(c(0.5, 0.5, 0, 0.5), "cm")
#         ),
#     p_beeswarm +
#         guides(color = "none") +
#         theme(
#             plot.margin = unit(c(0, 0.5, 0.5, 0.5), "cm")
#         ),
#     ncol = 1,
#     nrow = 2,
#     heights = c(2, 1),
#     guides = "collect"
# )
p_combined <- egg::ggarrange(
    p_scatter +
        labs(x = NULL) +
        guides(color = "none") +
        theme(
            axis.text.x = element_blank(),
            plot.margin = unit(c(0.5, 0.5, 0, 0.5), "cm")
        ),
    p_beeswarm +
        coord_fixed(ratio = .32) +
        theme(
            plot.margin = unit(c(0, 0.5, 0.5, 0.5), "cm"),
            legend.position = "bottom"
        ),
    nrow = 2,
    heights = c(2, 1.2)
)

ggsave(
    "plots/chembl_ic50_by_comparison_combined_new.pdf", p_combined,
    width = 8, height = 4.5
)
```

Now load the ChEMBL Kd only data and compare to KINOMEscan pseudo-IC50

Remove any ChEMBL Kd values that have relation that is not "=" unless
the relation is ">" or ">=" and the value is greater than 9000.


```{r}
chembl_kd_raw <- fread("data/kds_raw.csv.gz") %>%
  as_tibble()

chembl_kd <- chembl_kd_raw %>%
  mutate(across(standard_value, as.numeric)) %>%
  filter(
    standard_relation == "=" |
      (standard_relation %in% c(">=", ">") & standard_value > 9000)
  ) %>%
  power_inner_join(
    compound_mapping %>%
      filter(source == "chembl") %>%
      distinct(lspci_id, chembl_id_compound = external_id),
    by = "chembl_id_compound",
    check = check_specs(
      duplicate_keys_right = "warn",
      unmatched_keys_left = "warn"
    )
  ) %>%
  group_by(gene_symbol, lspci_id) %>%
  summarize(
    chembl_kd = quantile(standard_value, 0.25, na.rm = TRUE, names = FALSE),
    chembl_kd_relation = paste(sort(unique(standard_relation)), collapse = ""),
    .groups = "drop"
  ) %>%
  power_inner_join(
    kinomescan_kinase_info %>%
      distinct(hgnc_symbol, `DiscoveRx Gene Symbol`),
    by = c("gene_symbol" = "hgnc_symbol"),
    check = check_specs(
      duplicate_keys_right = "ignore",
      unmatched_keys_left = "warn"
    )
  )

chembl_kd %>%
  count(chembl_kd_relation)

pseudo_kd_kd_comp <- kinomescan_pseudo_kds %>%
  mutate(
    across(classification, \(x) factor(x, levels = names(class_colors)))
  ) %>%
  filter(
    dataset == "original_repeat_replaced",
    !exclude_target
  ) %>%
  inner_join(
    chembl_kd,
    by = c("lspci_id", "DiscoveRx Gene Symbol")
  )

p_scatter <- pseudo_kd_kd_comp %>%
  drop_na(pseudo_kd) %>%
  filter(
    # pseudo_ic50_relation == "=",
    pseudo_kd_concordant == "concordant",
    classification != "discordant"
  ) %>%
  ggplot(
    # aes(value, pseudo_ic50, color = classification, shape = pseudo_ic50_relation)
    aes(chembl_kd, pseudo_kd, color = classification)
  ) +
  scale_color_manual(values = class_colors) +
  geom_point(alpha = 0.5, shape = 16, size = 1) +
  # Limiting to between 1 and 10000 because barely any data outside this range
  scale_x_log10(labels = scales::label_number(big.mark = ""), limits = c(1, 10000)) +
  scale_y_log10(labels = scales::label_number(big.mark = "")) +
  coord_equal() +
  labs(
    x = "Kd (ChEMBL)",
    y = "Pseudo-Kd (KINOMEscan)"
  )

ggsave(
    "plots/chembl_kd_vs_kinomescan_pseudo_kd_new.pdf", p_scatter,
    width = 8, height = 4
)

p_beeswarm <- pseudo_kd_kd_comp %>%
  filter(
    pseudo_kd_concordant == "concordant",
    # pseudo_ic50_relation == "=",
    classification != "discordant"
  ) %>%
  mutate(
    classification = factor(
      classification,
      levels = names(class_colors) %>%
        rev()
    )
  ) %>%
  ggplot(
    aes(chembl_kd, classification, color = classification)
  ) +
  geom_quasirandom(orientation = "y", shape = 16, size = 1, alpha = 0.5, width = 0.3) +
  geom_boxplot(fill = NA, outlier.shape = NA) +
  scale_color_manual(values = class_colors) +
  guides(color = "none") +
  scale_x_log10(labels = scales::label_number(big.mark = ""), limits = c(1, 10000)) +
  scale_y_discrete(labels = class_labels) +
  theme(axis.text.y = element_markdown()) +
  labs(x = "Kd (ChEMBL)", y = NULL)

ggsave(
    "plots/chembl_kd_by_kinomescan_class_beeswarm.pdf", p_beeswarm,
    width = 7, height = 3
)

p_combined <- egg::ggarrange(
    p_scatter +
        labs(x = NULL) +
        guides(color = "none") +
        theme(
            axis.text.x = element_blank(),
            plot.margin = unit(c(0.5, 0.5, 0, 0.5), "cm")
        ),
    p_beeswarm +
        coord_fixed(ratio = .32) +
        theme(
            plot.margin = unit(c(0, 0.5, 0.5, 0.5), "cm"),
            legend.position = "bottom"
        ),
    nrow = 2,
    heights = c(2, 1.2)
)

ggsave(
    "plots/chembl_kd_by_comparison_combined_new.pdf", p_combined,
    width = 6, height = 4.5
)
```

Compute correlation coefficients between the ChEMBL IC50s / Kds and the KINOMEscan pseudo-IC50s.

Using Fisher r-to-z transformation to test for significant difference between the correlations.

```{r}
two_cor_test <- function(r1, r2, n1, n2) {
  (.5*log((1+r1)/(1-r1)) - .5*log((1+r2)/(1-r2))) /
    sqrt(1/(n1-3) + 1/(n2-3))
}

pseudo_kd_comp_both <- bind_rows(
  ic50 = pseudo_kd_comp %>%
    filter(
      pseudo_kd_concordant == "concordant",
      # pseudo_ic50_relation == "=",
      classification != "discordant"
    ) %>%
    select(lspci_id, name, `DiscoveRx Gene Symbol`, gene_symbol = hgnc_symbol, value, pseudo_kd, classification),
  kd = pseudo_kd_kd_comp %>%
    filter(
      pseudo_kd_concordant == "concordant",
      # pseudo_ic50_relation == "=",
      classification != "discordant"
    ) %>%
    select(lspci_id, name, `DiscoveRx Gene Symbol`, gene_symbol, value = chembl_kd, pseudo_kd, classification),
  .id = "chembl_type"
)

pseudo_kd_comp_cor_res <- pseudo_kd_comp_both %>%
  group_by(chembl_type) %>%
  summarize(
    res = cor.test(
      ~value + pseudo_kd, data = cur_data()
    ) %>%
      broom::tidy() %>%
      list(),
    n = n(),
    .groups = "drop"
  ) %>%
  unnest(res)

pseudo_kd_comp_cor_res

pseudo_kd_comp_cor_res %>%
  select(chembl_type, estimate, n) %>%
  pivot_wider(names_from = chembl_type, values_from = c(estimate, n)) %>%
  with(
    two_cor_test(
      estimate_ic50, estimate_kd,
      n_ic50, n_kd
    )
  )

```

Check false negatives and false positives. Any particular compounds or
targets that are consistently discordant?

Use Kd data

```{r}
pseudo_ic50_comp_false_class <- pseudo_ic50_comp_both %>%
  mutate(
    false_class = case_when(
      value > (10 * pseudo_ic50) ~ "false positive",
      value < (0.1 * pseudo_ic50) ~ "false negative",
      TRUE ~ "concordant"
    )
  )

pseudo_ic50_comp_false_class %>%
  count(chembl_type, false_class)

pseudo_ic50_comp_false_class_by_compound <- pseudo_ic50_comp_false_class %>%
  count(chembl_type, lspci_id, name, false_class) %>%
  group_by(chembl_type, lspci_id, name) %>%
  mutate(
    prop = n / sum(n)
  )

pseudo_ic50_comp_false_wide_by_compound <- pseudo_ic50_comp_false_class %>%
  count(chembl_type, lspci_id, name, false_class) %>%
  pivot_wider(names_from = false_class, values_from = n, values_fill = 0L)


pseudo_ic50_comp_false_wide_brms_input <- pseudo_ic50_comp_false_wide_by_compound %>%
  mutate(
    n_total = concordant + `false positive` + `false negative`
  ) %>%
  transmute(
    chembl_type,
    lspci_id, name,
    n_total,
    y = data.frame(
      concordant = concordant,
      `false positive` = `false positive`,
      `false negative` = `false negative`
    ) %>%
      as.matrix()
  )

library(brms)
mdl <- brm(
  y | trials(n_total) ~ 1 + (1 | name),
  data = pseudo_ic50_comp_false_wide_brms_input %>%
    filter(chembl_type == "kd"),
  family = multinomial(refcat = "concordant"),
  cores = 4,
  chains = 4
)

mdl2 <- brm(
  y | trials(n_total) ~ 1 + (1 | name) + chembl_type,
  data = pseudo_ic50_comp_false_wide_brms_input,
  family = multinomial(refcat = "concordant"),
  cores = 4,
  chains = 4
)

pseudo_ic50_comp_false_fisher_by_compound <- pseudo_ic50_comp_false_wide_by_compound %>%
  mutate(
    n_total = concordant + `false positive` + `false negative`
  ) %>%
  pivot_longer(
    c(`false positive`, `false negative`),
    names_to = "false_class",
    values_to = "n_class"
  ) %>%
  mutate(
    n_other = n_total - n_class
  ) %>%
  group_by(
    chembl_type, false_class
  ) %>%
  mutate(
    n_rest_class = sum(n_class) - n_class,
    n_rest_other = sum(n_other) - n_other
  ) %>%
  ungroup()



pseudo_ic50_comp_false_fisher_res_by_compound <- pseudo_ic50_comp_false_fisher_by_compound %>%
  group_by(chembl_type, lspci_id, name) %>%
  summarize(
    res = fisher.test(
      matrix(
        c(n_class, n_rest_class, n_other, n_rest_other),
        ncol = 2
      ),
      simulate = TRUE
    ) %>%
      broom::tidy() %>%
      list(),
    .groups = "drop"
  ) %>%
  unnest(res)

library(brms)
data(HairEyeColor)
haireye <- as.data.frame(HairEyeColor)
names(haireye) <- tolower(names(haireye))
names(haireye)[names(haireye) == "freq"] <- "number"
haireye
he_formula <- brmsformula(number ~ sex + hair + eye + sex:hair + sex:eye + hair:eye)
get_prior(he_formula, data=haireye, family='poisson')


pseudo_ic50_comp_false_class_by_kinase <- pseudo_ic50_comp_false_class %>%
  count(chembl_type, `DiscoveRx Gene Symbol`, false_class) %>%
  group_by(chembl_type, `DiscoveRx Gene Symbol`) %>%
  mutate(
    prop = n / sum(n)
  )
```

```{r}
pseudo_ic50_comp %>%
  filter(
    pseudo_ic50_concordant == "concordant",
    # pseudo_ic50_relation == "=",
    classification != "discordant"
  ) %>%
  group_by(classification) %>%
  mutate(ic50 = value) %>%
  summarize(
    across(
      c(pseudo_ic50, ic50),
      list(
        # average = ~10^(mean(log10(.x))),
        # ci = ~list(10^DescTools::MeanCI(log10(.x)))
        ci = ~list(10^quantile(log10(.x), c(.05, .5, .95), na.rm = TRUE))
      )
    ),
    .groups = "drop"
  ) %>%
  unnest_wider(ends_with("ci"), names_sep = "_")


pseudo_ic50_comp_left <- pseudo_ic50_class %>%
  group_by(
    lspci_id, hmsl_id, `Entrez Gene Symbol`, lspci_target_id
  ) %>%
  arrange(classification == "discordant", pseudo_ic50_relation != "=", pseudo_ic50) %>%
  slice_head(n = 1) %>%
  ungroup() %>%
  left_join(
    lsp_biochem,
    by = c("lspci_id", "lspci_target_id")
  )

pseudo_ic50_comp_left %>%
  mutate(
    chembl_is_covered = !is.na(value)
  ) %>%
  count(chembl_is_covered)


```

```{r ic50_single_dose_comp}
single_dose_trans <- single_dose %>%
  left_join(
    qpcr_misses %>%
      transmute(
        hmsl_id = `Compound Name`,
        `DiscoveRx Gene Symbol` = `Ambit Gene Symbol`,
        `Compound Conc`,
        qpcr_miss = TRUE
      ),
    by = c("hmsl_id", "DiscoveRx Gene Symbol", "Compound Concentration (nM)" = "Compound Conc")
  ) %>%
  mutate(
    log_inhibition = log10(`Percent Control` + 0.01),
    qpcr_miss = replace_na(qpcr_miss, FALSE)
  ) %>%
  as_tibble()


single_dose_comp <- single_dose_trans %>%
  power_left_join(
    lsp_targets %>%
      filter(organism == "Homo sapiens") %>%
      select(lspci_target_id, symbol),
    by = c("Entrez Gene Symbol" = "symbol"),
    check = check_specs(
      duplicate_keys_left = "ignore",
      duplicate_keys_right = "ignore",
      unmatched_keys_left = "warn",
      unmatched_keys_right = "ignore"
    )
  )  %>%
  power_left_join(
    kinomescan_classes %>%
      distinct(lspci_id, `DiscoveRx Gene Symbol`, classification) %>%
      mutate(across(classification, factor, levels = names(class_colors))),
    by = c("lspci_id", "DiscoveRx Gene Symbol"),
    check = check_specs(
      duplicate_keys_left = "ignore",
      duplicate_keys_right = "warn",
      unmatched_keys_left = "warn",
      unmatched_keys_right = "warn"
    )
  ) %>%
  inner_join(
    lsp_biochem,
    by = c("lspci_id", "lspci_target_id")
  )

p <- single_dose_comp %>%
  mutate(
    percent_control_zero = `Percent Control` == 0,
    percent_control_trans = if_else(
      percent_control_zero,
      0.5 * min(`Percent Control`[`Percent Control` > 0]),
      `Percent Control`
    ),
    value = pmin(value, 30000)
  ) %>%
  group_by(`Compound Concentration (nM)`) %>%
  filter(n() > 1000) %>%
  ungroup() %>%
  filter(
    !qpcr_miss
  ) %>%
  ggplot(
    aes(value, `Percent Control`)
  ) +
  # ggrastr::rasterize(geom_point(alpha = 0.5)) +
  ggrastr::rasterize(
    ggpointdensity::geom_pointdensity(alpha = 0.5, shape = 16, adjust = 2),
    dev = "ragg",
    dpi = 300
  ) +
  # scale_shape_manual(values = c(`TRUE` = 4, `FALSE` = 16)) +
  scale_x_log10(breaks = c(0.1, 1, 10, 100, 1000, 10000), labels = scales::label_number()) +
  scale_color_viridis_c(trans = scales::log10_trans(), option = "magma") +
  facet_grid(classification ~ `Compound Concentration (nM)`)

ggsave(
  "plots/single_dose_vs_chembl_ic50_all_doses.pdf", p,
  width = 10, height = 10
)

ps <- single_dose_comp %>%
  mutate(
    percent_control_zero = `Percent Control` == 0,
    percent_control_trans = if_else(
      percent_control_zero,
      0.5 * min(`Percent Control`[`Percent Control` > 0]),
      `Percent Control`
    ),
    value = pmin(value, 30000)
  ) %>%
  filter(
    !qpcr_miss
  ) %>%
  group_by(`Compound Concentration (nM)`) %>%
  summarize(
    plot = list(
      # ggplot(cur_data(), aes(value, percent_control_trans, shape = percent_control_zero)) +
      ggplot(cur_data(), aes(value, `Percent Control`, color = classification)) +
        geom_point(alpha = 0.5, shape = 16) +
        scale_color_manual(values = class_colors) +
        # scale_shape_manual(
        #   values = c(`TRUE` = 25, `FALSE` = 16)
        # ) +
        scale_x_log10(breaks = c(0.1, 1, 10, 100, 1000, 10000), labels = scales::label_number()) +
        # scale_y_log10() +
        guides(shape = "none") +
        labs(
          x = "IC50 ChEMBL (nM)",
          y = "% control Eurofins",
          title = paste("Compound concentration", cur_group()$`Compound Concentration (nM)`, "nM")
        )
    )
  )

pwalk(
  ps,
  function(`Compound Concentration (nM)`, plot, ...) {
    ggsave(
      paste0("plots/single_dose_vs_chembl_ic50_", `Compound Concentration (nM)`, "nM_with_class.pdf"),
      plot,
      width = 8, height = 6
    )
  }
)


ps <- single_dose_comp %>%
  mutate(
    percent_control_zero = `Percent Control` == 0,
    percent_control_trans = if_else(
      percent_control_zero,
      0.5 * min(`Percent Control`[`Percent Control` > 0]),
      `Percent Control`
    ),
    value = pmin(value, 30000)
  ) %>%
  filter(
    !qpcr_miss
  ) %>%
  group_by(`Compound Concentration (nM)`) %>%
  summarize(
    plot = list(
      # ggplot(cur_data(), aes(value, percent_control_trans, shape = percent_control_zero)) +
      ggplot(cur_data(), aes(value, `Percent Control`)) +
        geom_point(alpha = 0.5) +
        # scale_shape_manual(
        #   values = c(`TRUE` = 25, `FALSE` = 16)
        # ) +
        scale_x_log10(breaks = c(0.1, 1, 10, 100, 1000, 10000), labels = scales::label_number()) +
        # scale_y_log10() +
        guides(shape = "none") +
        labs(
          x = "IC50 ChEMBL (nM)",
          y = "% control Eurofins",
          title = paste("Compound concentration", cur_group()$`Compound Concentration (nM)`, "nM")
        )
    )
  )

pwalk(
  ps,
  function(`Compound Concentration (nM)`, plot, ...) {
    ggsave(
      paste0("plots/single_dose_vs_chembl_ic50_", `Compound Concentration (nM)`, "nM.pdf"),
      plot,
      width = 8, height = 8
    )
  }
)
```

Testing if there are any kinases or compounds that are particularly discordant
compared to the ChEBML data.

First, establishing the Percent Control thresholds necessary to ensure a 95%
probability that the ChEMBL IC50 is below 5000 nM.

To do that, first plot for each possible Percent Control threshold what proportion
of IC50s is below 5000 nM.


```{r}
percent_control_threshold_plotting_data <- single_dose_comp %>%
  group_by(`Compound Concentration (nM)`) %>%
  filter(n() > 1000) %>%
  ungroup() %>%
  filter(
    classification != "discordant",
    !qpcr_miss
  ) %>%
  # crossing(ic_50_threshold = 10**seq(2, 4, length.out = 6)) %>%
  crossing(ic_50_threshold = c(100, 200, 500, 1000, 5000, 10000)) %>%
  group_by(`Compound Concentration (nM)`, ic_50_threshold) %>%
  arrange(`Percent Control`) %>%
  mutate(
    cumsum_below_threshold = cumsum(value < ic_50_threshold),
    frac_below_threshold = cumsum_below_threshold / seq_len(n())
  ) %>%
  ungroup() %>%
  group_by(`Compound Concentration (nM)`, `Percent Control`, ic_50_threshold) %>%
  slice_tail(n = 1) %>%
  ungroup()

p <- percent_control_threshold_plotting_data %>%
  mutate(across(c(ic_50_threshold, `Compound Concentration (nM)`), ~as.character(.x) %>% fct_inseq(ordered = TRUE))) %>%
  ggplot(aes(`Percent Control`, frac_below_threshold, color = `Compound Concentration (nM)`)) +
  geom_line() +
  # geom_point() +
  facet_wrap(~ic_50_threshold) +
  labs(y = "Fraction of measurements below IC50 threshold")

ggsave(
  "plots/percent_control_below_ic50_threshold_no_discordant.pdf", p,
  width = 8, height = 4
)

```


```{r}
pseudo_ic50_comp_stats <- pseudo_ic50_comp %>%
  drop_na(pseudo_ic50) %>%
  filter(classification != "discordant") %>%
  mutate(
      log10_ic50 = log10(value),
      log10_pseudo_ic50 = log10(pseudo_ic50),
      log_ic50_diff = log10_ic50 - log10_pseudo_ic50,
      diff_greater_than_1 = abs(log_ic50_diff) > 1,
      difference_class = case_when(
          log_ic50_diff > 1 ~ "FP",
          log_ic50_diff < -1 ~ "FN",
          TRUE ~ "concordant"
      )
  )
```


```{r}
stats_per_group <- function(grouped_df) {
    grouped_df %>%
    summarize(
        n = n(),
        n_binding_class = sum(classification == "binding"),
        frac_binding_class = n_binding_class / n,
        n_diff_greater_than_1 = sum(diff_greater_than_1, na.rm = TRUE),
        frac_diff_greater_than_1 = n_diff_greater_than_1 / n,
        mean_log_ic50_diff = mean(log_ic50_diff, na.rm = TRUE),
        sd_log_ic50_diff = sd(log_ic50_diff, na.rm = TRUE),
        n_false_positives = sum(difference_class == "FP", na.rm = TRUE),
        n_false_negatives = sum(difference_class == "FN", na.rm = TRUE),
        frac_false_positives = n_false_positives / n,
        frac_false_negatives = n_false_negatives / n,
        .groups = "drop"
    )
}

pseudo_ic50_comp_stats_per_compound <- pseudo_ic50_comp_stats %>%
    group_by(lspci_id, name) %>%
    filter(n() > 10) %>%
    stats_per_group()
```

```{r}
library(ggrepel)
p <- pseudo_ic50_comp_stats_per_compound %>%
  mutate(
    outside_prediction_interval = {
      mdl <- lm(
        frac_diff_greater_than_1 ~ frac_binding_class,
        data = .
      )
      pred_interval <- predict(mdl, interval = "prediction", level = 0.8)
      frac_diff_greater_than_1 < pred_interval[, "lwr"] | frac_diff_greater_than_1 > pred_interval[, "upr"]
    }
  ) %>%
  ggplot(
    aes(
      frac_binding_class, frac_diff_greater_than_1,
      text = name
    )
  ) +
    geom_point(shape = 16) +
    geom_text_repel(
      aes(label = name),
      data = ~.x %>%
        mutate(name = if_else(!outside_prediction_interval, "", name)),
      # min.segment.length = 0,
      max.overlaps = Inf,
    ) +
    guides(color = "none") +
    labs(
        x = "Fraction of 'binding' dose-response curves",
        y = "Fraction of pseudo IC50s with >10x difference",
        title = "Per compound"
    )


plotly::ggplotly(p)

ggsave(
    "plots/eurofins_ic50_vs_kinomescan_pseudo_ic50_differences_per_compound.pdf", p,
    width = 10, height = 7
)
```

Now do the same per kinase

```{r}

pseudo_ic50_comp_stats_per_kinase <- pseudo_ic50_comp_stats %>%
    group_by(`DiscoveRx Gene Symbol`) %>%
    filter(n() > 10) %>%
    stats_per_group() %>%
    inner_join(
      kinomescan_kinase_info
    )
```

```{r}
lm(
  frac_diff_greater_than_1 ~ frac_binding_class,
  data = pseudo_ic50_comp_stats_per_kinase
) %>%
  summary()

cor.test(
  ~frac_diff_greater_than_1 + frac_binding_class,
  data = pseudo_ic50_comp_stats_per_kinase
)

p <- pseudo_ic50_comp_stats_per_kinase %>%
  mutate(
    outside_prediction_interval = {
      mdl <- lm(
        frac_diff_greater_than_1 ~ frac_binding_class,
        data = .
      )
      pred_interval <- predict(mdl, interval = "prediction", level = 0.8)
      frac_diff_greater_than_1 < pred_interval[, "lwr"] | frac_diff_greater_than_1 > pred_interval[, "upr"]
    }
  ) %>%
  ggplot(
    aes(
      frac_binding_class, frac_diff_greater_than_1,
      color = `Kinase Form` == "Wild Type",
      text = `DiscoveRx Gene Symbol`
    )
  ) +
    geom_point(shape = 16) +
    geom_text_repel(
      aes(label = `DiscoveRx Gene Symbol`),
      data = ~.x %>%
        mutate(`DiscoveRx Gene Symbol` = if_else(!outside_prediction_interval, "", `DiscoveRx Gene Symbol`)),
      # min.segment.length = 0,
      max.overlaps = Inf,
      box.padding = 0.5,
      max.iter = 50000, max.time = 50,
      force = 1.5,
      seed = 42
    ) +
    guides(color = "none") +
    labs(
        x = "Fraction of 'binding' dose-response curves",
        y = "Fraction of pseudo IC50s with >10x difference",
        title = "Per kinase"
    )

plotly::ggplotly(p)

ggsave(
    "plots/eurofins_ic50_vs_kinomescan_pseudo_ic50_differences_per_kinase.pdf", p,
    width = 10, height = 7
)
```

Establish confidence intervals for IC50 values given the Eurofins data

Doing it in bins of %control values as well as cumulatively

```{r}
single_dose_comp_binned <- single_dose_comp %>%
  mutate(
    percent_control_bin = cut(
      `Percent Control`,
      breaks = c(-Inf, seq(from = 10, to = 90, by = 10), Inf)
    ),
    ic_50_bin = cut(
      value,
      breaks = c(-Inf, 10**seq(-1, 4), Inf)
    )
  )

percent_control_ic50_ci <- single_dose_comp_binned %>%
  group_by(`Compound Concentration (nM)`) %>%
  filter(n() > 1000) %>%
  ungroup() %>%
  filter(
    classification != "discordant",
    !qpcr_miss
  ) %>%
  group_by(`Compound Concentration (nM)`, percent_control_bin) %>%
  summarize(
    n = n(),
    ci = quantile(value, probs = c(0.05, .5, 0.95)) %>%
      enframe("quantile", "value") %>%
      # pivot_wider(names_from = quantile, values_from = value, names_prefix = "ci_") %>%
      list(),
    .groups = "drop"
  ) %>%
  unnest(ci)

p <- percent_control_ic50_ci %>%
  ggplot(
    aes(percent_control_bin, value, color = quantile, group = quantile)
  ) +
  geom_point() +
  geom_line() +
  facet_wrap(~`Compound Concentration (nM)`) +
  scale_y_log10()

percent_control_ic50_ci_ic50 <- single_dose_comp_binned %>%
  group_by(`Compound Concentration (nM)`) %>%
  filter(n() > 1000) %>%
  ungroup() %>%
  filter(
    classification != "discordant",
    !qpcr_miss
  ) %>%
  group_by(`Compound Concentration (nM)`, ic_50_bin) %>%
  summarize(
    n = n(),
    ci = quantile(`Percent Control`, probs = c(0.05, .25, .5, .75, 0.95)) %>%
      enframe("quantile", "value") %>%
      # pivot_wider(names_from = quantile, values_from = value, names_prefix = "ci_") %>%
      list(),
    .groups = "drop"
  ) %>%
  unnest(ci)

p <- percent_control_ic50_ci_ic50 %>%
  ggplot(
    aes(ic_50_bin, value, color = quantile, group = quantile)
  ) +
  geom_point() +
  geom_line() +
  facet_wrap(~`Compound Concentration (nM)`) +
  scale_y_log10()

p <- single_dose_comp_binned %>%
  ggplot(
    aes(ic_50_bin, value)
  ) +
  ggrastr::rasterize(geom_beeswarm()) +
  facet_wrap(~`Compound Concentration (nM)`) +
  scale_y_log10()


p <- single_dose_comp_binned %>%
  group_by(`Compound Concentration (nM)`) %>%
  filter(n() > 1000) %>%
  ungroup() %>%
  mutate(
    across(
      `Compound Concentration (nM)`,
      ~ordered(
        .x,
        levels = sort(unique(.x)),
        labels = paste0(sort(unique(.x)), " nM")
      )
    )
  ) %>%
  filter(
    classification != "discordant",
    !qpcr_miss
  ) %>%
  ggplot(
    aes(
      `Percent Control`, value,
      color = after_stat(quantile) %>% {
        factor(
          .,
          levels = sort(unique(.)),
          labels = scales::label_percent()(sort(unique(.)))
        )
      }
      # color = fct_inseq(as.character(after_stat(quantile)), ordered = TRUE)
    )
  ) +
  geom_quantile(
    method = "rqss",
    quantiles = c(.05, .25, .5, .75, .95),
    lambda = 10
  ) +
  # geom_point() +
  # geom_smooth(method = "lm", se = FALSE) +
  scale_y_log10(
    breaks = 10**(0:4)
  ) +
  scale_color_manual(
    values = c(
      "5%" = "red",
      "25%" = "orange",
      "50%" = "black",
      "75%" = "orange",
      "95%" = "red"
    )
  ) +
  labs(
    x = "% control (Eurofins)",
    y = "IC50 (ChEMBL)",
    color = "Quantile"
  ) +
  facet_wrap(~`Compound Concentration (nM)`)

ggsave(
  "plots/percent_control_ic50_ci_quantile_plot.pdf",
  p, width = 8, height = 6
)

p <- single_dose_comp_binned %>%
  group_by(`Compound Concentration (nM)`) %>%
  filter(n() > 1000) %>%
  ungroup() %>%
  mutate(
    across(
      `Compound Concentration (nM)`,
      ~ordered(
        .x,
        levels = sort(unique(.x)),
        labels = paste0(sort(unique(.x)), " nM")
      )
    )
  ) %>%
  filter(
    classification != "discordant",
    !qpcr_miss
  ) %>%
  ggplot(
    aes(
      value, `Percent Control`,
      color = after_stat(quantile) %>% {
        factor(
          .,
          levels = sort(unique(.)),
          labels = scales::label_percent()(sort(unique(.)))
        )
      }
      # color = fct_inseq(as.character(after_stat(quantile)), ordered = TRUE)
    )
  ) +
  geom_quantile(
    method = "rqss",
    quantiles = c(.05, .25, .5, .75, .95),
    lambda = 10
  ) +
  # geom_point() +
  # geom_smooth(method = "lm", se = FALSE) +
  scale_x_log10(
    breaks = 10**(0:4)
  ) +
  scale_color_manual(
    values = c(
      "5%" = "red",
      "25%" = "orange",
      "50%" = "black",
      "75%" = "orange",
      "95%" = "red"
    )
  ) +
  labs(
    x = "IC50 (ChEMBL)",
    y = "% control (Eurofins)",
    color = "Quantile"
  ) +
  facet_wrap(~`Compound Concentration (nM)`)

ggsave(
  "plots/percent_control_ic50_ci_quantile_plot_reverse.pdf",
  p, width = 8, height = 6
)
```

```{r}
IC50_THRESHOLDS <- c(100, 1000, 5000)
library(ROCit)

single_dose_comp_binned_filtered <- single_dose_comp_binned %>%
  group_by(`Compound Concentration (nM)`) %>%
  filter(n() > 1000) %>%
  ungroup() %>%
  mutate(
    across(
      `Compound Concentration (nM)`,
      ~ordered(
        .x,
        levels = sort(unique(.x)),
        labels = paste0(sort(unique(.x)), " nM")
      )
    )
  ) %>%
  filter(
    classification != "discordant",
    !qpcr_miss
  )

single_dose_comp_rocs <- single_dose_comp_binned_filtered %>%
  crossing(
    ic50_threshold = IC50_THRESHOLDS
  ) %>%
  group_by(`Compound Concentration (nM)`, ic50_threshold) %>%
  summarize(
    roc = rocit(
      100 - `Percent Control`,
      value < cur_group()$ic50_threshold,
      method = "empirical"
    ) %>%
      list(),
    .groups = "drop"
  ) %>%
  mutate(
    # roc_ci = map(roc, ciROC),
    roc_measures = map(
      roc,
      ~measureit(.x, measure = c("SENS", "SPEC", "PREC")) %>%
        unclass() %>%
        as_tibble()
    )
  )


single_dose_comp_roc_values <- single_dose_comp_rocs %>%
  transmute(
    `Compound Concentration (nM)`,
    ic50_threshold = fct_inorder(
      paste(ic50_threshold, "nM")
    ),
    roc_values = map(roc, plot)
  )

single_dose_comp_roc_values_optimal <- single_dose_comp_roc_values %>%
  mutate(
    across(
      roc_values,
      map,
      ~.x$`optimal Youden Index point` %>%
        as.list() %>%
        as_tibble()
    )
  ) %>%
  unnest(roc_values) %>%
  transmute(
    `Compound Concentration (nM)`,
    ic50_threshold,
    `Cutoff` = 100 - cutoff,
    `FPR/1-Specificity` = FPR,
    `TPR/Sensitivity/Recall` = TPR
  )


pwalk(
  single_dose_comp_rocs,
  function(`Compound Concentration (nM)`, ic50_threshold, roc, ...) {
    withr::with_pdf(
      paste0("plots/percent_control_ic50_roc_threshold_", ic50_threshold, "_", `Compound Concentration (nM)`, ".pdf"),
      plot(roc)
    )
  }
)

pwalk(
  single_dose_comp_rocs,
  function(`Compound Concentration (nM)`, ic50_threshold, roc, ...) {
    withr::with_pdf(
      paste0("plots/percent_control_ic50_ks_threshold_", ic50_threshold, "_", `Compound Concentration (nM)`, ".pdf"),
      ksplot(roc)
    )
  }
)

```

Table with FP, FN, AUC for each compound concentration and threshold
highlighted best and worst screening concentration for each

```{r}
confusion_tables <- single_dose_comp_binned_filtered %>%
  inner_join(
    single_dose_comp_roc_values_optimal %>%
      select(
        `Compound Concentration (nM)`,
        ic50_threshold,
        `Cutoff`
      ),
    by = "Compound Concentration (nM)",
    relationship = "many-to-many"
  ) %>%
  group_by(`Compound Concentration (nM)`, ic50_threshold) %>%
  summarize(
    confusion_table = list(
      table(
        chembl_ic50_below = value < as.double(str_replace(ic50_threshold, " nM", "")),
        eurofins_below = `Percent Control` < `Cutoff`
      )
    ),
    .groups = "drop"
  )

library(cvms)

# p <- plot_confusion_matrix(
#   confusion_tables$confusion_table[[1]] %>%
#   as_tibble(), 
#   target_col = "chembl_ic50_below", 
#   prediction_col = "eurofins_below",
#   counts_col = "n"
# )
# ggsave(
#   "test.pdf", p, width = 2.5, height = 2.5
# )

dir.create("plots/percent_control_ic50_roc_confusion_tables", showWarnings = FALSE)
pwalk(
  confusion_tables,
  function(`Compound Concentration (nM)`, ic50_threshold, confusion_table, ...) {
    p <- plot_confusion_matrix(
      confusion_table %>%
        as_tibble(), 
      target_col = "chembl_ic50_below", 
      prediction_col = "eurofins_below",
      counts_col = "n"
    ) +
      labs(
        subtitle = paste("Predicting IC50 <", ic50_threshold, "\nat", `Compound Concentration (nM)`)
      ) +
      theme(
        plot.subtitle = element_text(size = 8)
      )
    ggsave(
      paste0("plots/percent_control_ic50_roc_confusion_tables/confusion_table_", ic50_threshold, "_", `Compound Concentration (nM)`, ".pdf"),
      p, width = 2.5, height = 3
    )
  }
)



```

Plot precision-recall curves

```{r}

library(ggrepel)
single_dose_comp_pr_plots <- single_dose_comp_rocs %>%
  transmute(
    `Compound Concentration (nM)`,
    ic50_threshold = fct_inorder(
      paste(ic50_threshold, "nM")
    ),
    roc_measures
  ) %>%
  unnest(roc_measures) %>%
  mutate(across(Cutoff, ~100 - .x)) %>%
  ggplot(
    aes(
      SENS, PREC,
      color = fct_inorder(
        as.character(`Compound Concentration (nM)`)
      )
    )
  ) +
  geom_line() +
  geom_point(
    data = ~.x %>%
      filter(
        Cutoff %in% seq(from = 0, to = 100, by = 25)
      ) %>%
      bind_rows(
        inner_join(
          .x,
          single_dose_comp_roc_values_optimal
        )
      ),
    show.legend = FALSE
  ) +
  geom_text_repel(
    aes(label = Cutoff),
    data = ~.x %>%
      mutate(
        Cutoff = if_else(
          Cutoff %in% seq(from = 25, to = 75, by = 25),
          as.character(Cutoff),
          ""
        )
      ) %>%
      anti_join(
        single_dose_comp_roc_values_optimal %>%
          mutate(across(Cutoff, as.character))
      ) %>%
      bind_rows(
        inner_join(
          .x,
          single_dose_comp_roc_values_optimal
        ) %>%
        mutate(
          across(Cutoff, ~paste0(.x, "*"))
        )
      ),
    box.padding = 0.5,
    show.legend = FALSE,
    max.iter = 50000,
    max.time = 2,
    force = 3,
    seed = 42,
    max.overlaps = Inf
  ) +
  labs(
    x = "Recall TP / (TP + FN)",
    y = "Precision TP / (TP + FP)",
    color = "Assay concentration",
    title = "Precision-recall curves for prediction of IC50 < threshold"
  ) +
  facet_wrap(~ic50_threshold) +
  coord_cartesian(xlim = c(0, 1.1))

ggsave(
  "plots/percent_control_ic50_pr.pdf",
  single_dose_comp_pr_plots,
  width = 10, height = 4
)

single_dose_comp_roc_plot <- single_dose_comp_roc_values %>%
  mutate(
    across(
      roc_values,
      map,
      ~.x[c("FPR/1-Specificity", "TPR/Sensitivity/Recall", "Cutoff")] %>%
        as_tibble()
    )
  ) %>%
  unnest(roc_values) %>%
  mutate(across(Cutoff, ~100 - .x)) %>%
  ggplot(
    aes(
      `FPR/1-Specificity`, `TPR/Sensitivity/Recall`,
      color = fct_inorder(
        as.character(`Compound Concentration (nM)`)
      )
    )
  ) +
  geom_line() +
  geom_point(
    data = ~.x %>%
      filter(
        Cutoff %in% seq(from = 0, to = 100, by = 25)
      ) %>%
      bind_rows(
        single_dose_comp_roc_values_optimal
      ),
    show.legend = FALSE
  ) +
  geom_text_repel(
    aes(label = Cutoff),
    data = ~.x %>%
      mutate(
        Cutoff = if_else(
          Cutoff %in% seq(from = 25, to = 75, by = 25),
          as.character(Cutoff),
          ""
        )
      ) %>%
      # If optimal value is equal to one of the plotted
      # values don't repeat it
      anti_join(
        single_dose_comp_roc_values_optimal %>%
          mutate(across(Cutoff, as.character))
      ) %>%
      bind_rows(
        single_dose_comp_roc_values_optimal %>%
          mutate(across(Cutoff, ~paste0(.x, "*")))
      ),
    # point.padding = unit(20, "pt"),
    # min.segment.length = 0,
    box.padding = 0.5,
    show.legend = FALSE,
    max.iter = 50000,
    max.time = 2,
    force = 3,
    seed = 42,
    max.overlaps = Inf
  ) +
  labs(
    x = "1 - Specificity TN / (TN + FP)",
    y = "Sensitivity TP / (TP + FN)",
    color = "Assay concentration",
    title = "ROC curve for predicting an IC50 below threshold"
  ) +
  facet_wrap(~ic50_threshold) +
  coord_cartesian(xlim = c(-0.1, 1), ylim = c(0, 1.1))

ggsave(
  "plots/percent_control_ic50_roc.pdf",
  single_dose_comp_roc_plot,
  width = 12, height = 4
)
```

