
## Scrape Eurofins kinase assay datasheets

Eurofins doesn't make information about the kinases they measure available in a
machine-readable format. This is a quick and dirty script to scrape the
datasheets for each target and extract whether the target is mutated,
phosphorylated, and what form the kinase is assayed in. E.g. only a
domain or the entire kinase.

```{r}
library(tidyverse)
library(rvest)
library(qs)
library(synExtra)

synapser::synLogin()
```

```{r}
assay_index <- read_html("https://www.discoverx.com/services/drug-discovery-development-services/kinase-profiling/kinomescan#kinase-assay-list")

assay_table_node <- assay_index %>%
    html_elements("table") %>%
    magrittr::extract2(3)

assay_table_raw <- html_table(assay_table_node, header = TRUE)
```


```{r}
assay_links <- assay_table_node %>%
    html_elements("td a") %>%
    html_attr("href") %>%
    str_subset("kinase-data-sheets") %>%
    str_replace(fixed("../../.."), "https://www.discoverx.com")

length(assay_links) == nrow(assay_table_raw)

assay_datasheet_nodes <- map(assay_links, read_html)


```

Fetch each datasheet and extract the assay information. Put reference compound
information in a separate column (not particularly interesting).

```{r}
extract_assay_info <- function(datasheet_node) {
    vals_raw <- tibble(
        name = datasheet_node %>%
            html_nodes("dt") %>%
            html_text(),
        value = datasheet_node %>%
            html_nodes("dd") %>%
            html_text()
    )
    if ("Compound" %in% vals_raw$name) {
        cmpd_info <- vals_raw %>%
            slice(which(vals_raw$name == "Compound") + 1:nrow(vals_raw)) %>%
            magrittr::set_colnames(c("Compound", "Kd"))
        vals <- vals_raw %>%
            slice(1:(which(vals_raw$name == "Compound") - 1)) %>%
            pivot_wider(names_from = name, values_from = value) %>%
            mutate(Compounds = list(cmpd_info))
    } else
        vals <- vals_raw %>%
        pivot_wider(names_from = name, values_from = value)
    vals
}

assay_datasheets_raw <- assay_datasheet_nodes %>%
    map(extract_assay_info)
```

Include Ensembl and Entrez gene IDs and UniProt IDs for each kinase.

```{r}
library(biomaRt)
mart <- biomaRt::useMart("ensembl", dataset = "hsapiens_gene_ensembl")
gene_symbol_to_id_map <- biomaRt::getBM(
    attributes = c("ensembl_gene_id", "external_gene_name", "entrezgene_id", "chromosome_name"),
    filters = "external_gene_name",
    values = assay_table_raw$`Entrez Gene Symbol`,
    mart = mart
)
setdiff(assay_table_raw$`Entrez Gene Symbol`, gene_symbol_to_id_map$external_gene_name)

# figure out ambiguous mappings
gene_symbol_to_id_map %>%
    group_by(external_gene_name) %>%
    filter(n() > 1) %>%
    arrange(external_gene_name) %>%
    print(n = Inf)

gene_symbol_to_id_map_disam <- gene_symbol_to_id_map %>%
    filter(
        # ensembl gene id hits for some are on alternative assemblies
        chromosome_name %in% c(as.character(1:22), "X", "Y"),
        # this entrez gene hit is a BUB1B-PAK6 readthrough
        !entrezgene_id %in% c(106821730)
    )
setdiff(assay_table_raw$`Entrez Gene Symbol`, gene_symbol_to_id_map_disam$external_gene_name) %>%
    sort()

manual_map_old_symbols <- tribble(
  ~`Entrez Gene Symbol`, ~entrezgene_id, ~ensembl_gene_id,
    "ADCK4", 79934, "ENSG00000123815",
    "ADRBK1", 156, "ENSG00000173020",
    "ADRBK2", 157, "ENSG00000100077",
    "CABC1", 56997, "ENSG00000163050",
    "CDC2L2", 728642, "ENSG00000008128",
    "CDPK1", 812762, NA_character_, # Plasmodium
    "GSG2", 83903, "ENSG00000177602",
    "ICK", 22858, "ENSG00000112144",
    "KIAA0999", 23387, "ENSG00000160584",
    "MAL13P1.279", 813841, NA_character_, # Plasmodium
    "MGC42105", 167359, "ENSG00000177453",
    "MST4", 51765, "ENSG00000134602",
    "PAK7", 57144, "ENSG00000101349",
    "pknB", 887072, NA_character_, # M. tuberculosis
    "SgK110", 100130827, "ENSG00000231274",
    "ZAK", 51776, "ENSG00000091436"
)
gene_symbol_to_id_map_final <- gene_symbol_to_id_map_disam %>%
    distinct(`Entrez Gene Symbol` = external_gene_name, entrezgene_id, ensembl_gene_id) %>%
    bind_rows(manual_map_old_symbols)
setdiff(assay_table_raw$`Entrez Gene Symbol`, gene_symbol_to_id_map_final$`Entrez Gene Symbol`) %>%
    sort()

```

```{r}
assay_datasheets <- assay_table_raw %>%
    bind_cols(
        bind_rows(assay_datasheets_raw)
    ) %>%
    left_join(
        gene_symbol_to_id_map_final
    ) %>%
    dplyr::select(
        `DiscoveRx Gene Symbol` = `KGS ▲`,
        `Entrez Gene Symbol`,
        `Kinase Name`,
        entrezgene_id,
        ensembl_gene_id,
        everything()
    )
```


```{r}
write_csv(
    assay_datasheets %>%
     dplyr::select(where(negate(is.list))),
    "eurofins_kinase_info.csv"
)
qsave(
    assay_datasheets,
    "eurofins_kinase_info.qs"
)

synStoreMany(
    c("eurofins_kinase_info.csv", "eurofins_kinase_info.qs"),
    parent = "syn18502717",
    activity = synapser::Activity(
        used = c(
            "https://www.discoverx.com/services/drug-discovery-development-services/kinase-profiling/kinomescan#kinase-assay-list"
        ),
        executed = "https://github.com/labsyspharm/okl-analysis/blob/main/eurofins_assays.Rmd"
    )
)
```
