
```{r}
library(tidyverse)
library(synExtra)
library(fst)
library(data.table)
library(powerjoin)

synapser::synLogin()

syn <- synDownloader("~/data", .cache = TRUE)
```

```{r}
kinomescan_pseudo_ic50s <- syn("syn51080578") %>%
  read_csv()

kinomescan_classes <- syn("syn51080579") %>%
  read_csv()

kinomescan_kinase_info <- syn("syn51286743") %>%
  read_csv()

single_dose <- syn("syn26486828") %>%
  fread()

download.file("https://github.com/sorgerlab/famplex/raw/43732660e42fd1746c2d1e61552bda0a90cbd664/relations.csv", destfile = "~/data/famplex_relations.csv")
famplex_relations <- read_csv("~/data/famplex_relations.csv", col_names = c("namespace", "id", "relation", "family_namespace", "family"))
```

Extract the kinases from the famplex relations.

Use list of preselected family kinases

```{r}
famplex_kinases <- famplex_relations %>%
  filter(
    relation == "isa",
    id %in% kinomescan_kinase_info$hgnc_symbol,
    family %in% c(
      "AKT", "AMPK_alpha", "CDK", "CSNK1",
      "ERK", "FGFR", "GSK3", "PDGFR", "ERBB",
      "JAK", "P90RSK", "MAP3K", "TAOK", "MAP2K",
      "PAK", "ROCK", "p38", "ADRBK", "DDR", "EPHA",
      "EPHB", "GRK", "IKK_family", "JNK", "NTRK",
      "PKC", "SRC", "VEGFR", "PIM", "SIK", "DYRK",
      "RSTK1", "RSTK2", "DAPK", "CLK", "RIPK", "PKN", "LATS"
    )
  ) %>%
  distinct(hgnc_symbol = id, family) %>%
  # manually add HIP kinases
  bind_rows(
    tibble(
      hgnc_symbol = c("HIPK1", "HIPK2", "HIPK3", "HIPK4"),
      family = "HIP"
    )
  )
```

Clean pseudo IC50s by retaining only concordant values and dose-response curves
that are not discordant.

```{r}
pseudo_ic50s_clean <- kinomescan_pseudo_ic50s %>%
  # filter(
  #   pseudo_ic50_concordant == "concordant"
  # ) %>%
  inner_join(
    kinomescan_classes %>%
      filter(
        classification != "discordant"
      ) %>%
      select(lspci_id, `DiscoveRx Gene Symbol`, classification)
  ) %>%
  # Remove any mutants
  inner_join(
    kinomescan_kinase_info %>%
      filter(`Kinase Form` == "Wild Type") %>%
      select(`DiscoveRx Gene Symbol`, hgnc_symbol)
  ) %>%
  # Select kinase form (complex or PTM) with the strongest inhibition
  filter(
    `DiscoveRx Gene Symbol` %in% {
      group_by(., `DiscoveRx Gene Symbol`, hgnc_symbol) %>%
        summarize(
          frac_binding = sum(classification == "binding") / n(),
          .groups = "drop"
        ) %>%
        group_by(hgnc_symbol) %>%
        arrange(desc(frac_binding)) %>%
        slice_head(n = 1) %>%
        pull(`DiscoveRx Gene Symbol`)
    }
  ) %>%
  # group_by(lspci_id, name, hgnc_symbol, `Entrez Gene Symbol`) %>%
  # arrange(classification == "binding", pseudo_ic50) %>%
  # slice_head(n = 1) %>%
  # ungroup() %>%
  power_inner_join(
    famplex_kinases %>%
      select(hgnc_symbol, family),
    check = check_specs(
      duplicate_keys_right = "warn"
    )
  )
```

Compute number of compounds inhibiting each kinase

```{r}
kinase_stats <- pseudo_ic50s_clean %>%
  group_by(hgnc_symbol, family, `DiscoveRx Gene Symbol`) %>%
  summarize(
    n_binding = sum(classification == "binding"),
    .groups = "drop"
  ) %>%
  arrange(desc(n_binding))
```

Group kinases by family and compute some summary stats for each family member,
like the number of compounds inhibiting it

```{r}
kinase_stats_by_family <- kinase_stats %>%
  mutate(
    across(c(hgnc_symbol, `DiscoveRx Gene Symbol`), fct_inorder),
    family = factor(
      family,
      levels = group_by(., family) %>%
        summarize(max_spread = max(n_binding) - min(n_binding), .groups = "drop") %>%
        arrange(desc(max_spread)) %>%
        pull(family)
    )
  )

library(ggpattern)
library(ggrastr)
p <- kinase_stats_by_family %>%
  inner_join(
    kinomescan_kinase_info %>%
      select(`DiscoveRx Gene Symbol`, `Expression System`, `Kinase Construct`) %>%
      mutate(across(`DiscoveRx Gene Symbol`, factor, levels = kinase_stats$`DiscoveRx Gene Symbol`))
  ) %>%
  mutate(
    across(c(`Kinase Construct`, `Expression System`), factor),
  ) %>%
  ggplot(
    aes(
      x = `DiscoveRx Gene Symbol`,
      y = n_binding,
      fill = `Expression System`,
      pattern = `Kinase Construct`
    )
  ) +
    coord_flip() +
    # geom_col() +
    geom_col_pattern(color = "black") +
    scale_pattern_manual(values = c("Full Length" = "none", "Partial Length" = "circle")) +
    facet_wrap(~family, scales = "free") +
    theme_minimal() +
    theme(legend.key.size = unit(1, "cm")) +
    labs(x = "Kinase", y = "Number of compounds inhibiting kinase")

ggsave(
  "plots/kinase_families_n_binding.pdf", p,
  width = 16,
  height = 12,
)
```

```{r}
library(seriation)
co_cluster_mats <- function(mats) {
  set.seed(42)
  dms <- map(mats, dist, method = "euclidean")
  dm <- reduce(dms, `+`)
  cr <- hclust(dm, method = "average") %>%
    reorder(dm, method = "olo")
  cr
}

cluster_mat <- function(mat) {
  set.seed(42)
  dr <- dist(mat, method = "euclidean")
  cr <- hclust(dr, method = "average") %>%
    reorder(dr, method = "olo")
  cr
}

library(pheatmap)
plot_mat <- function(mat, cr = NULL, cc = NULL) {
  viridis_map <- viridis::plasma(100)
  pheatmap::pheatmap(
    mat,
    show_rownames = FALSE, show_colnames = TRUE,
    color = rev(viridis_map),
    cluster_rows = if (is.null(cr)) cluster_mat(mat) else cr,
    cluster_cols = if (is.null(cc)) cluster_mat(t(mat)) else cc,
    treeheight_row = 20, treeheight_col = 20,
    silent = TRUE
  )
}

generate_mat <- function(targets) {
  all_doses <- single_dose %>%
    filter(
      `DiscoveRx Gene Symbol` %in% targets
    ) %>%
    select(
      lspci_id, `DiscoveRx Gene Symbol`, `Compound Concentration (nM)`, `Percent Control`
    ) %>%
    arrange(lspci_id, `DiscoveRx Gene Symbol`, `Compound Concentration (nM)`) %>%
    group_by(`Compound Concentration (nM)`) %>%
    filter(n() > 10) %>%
    ungroup()
  measured_in_all <- all_doses %>%
    group_by(lspci_id) %>%
    summarize(
      n = n(),
      .groups = "drop"
    ) %>%
    filter(n == max(n)) %>%
    pull(lspci_id)
  all_doses_mats <- all_doses %>%
    mutate(across(`Compound Concentration (nM)`, ~factor(as.character(.x), levels = as.character(sort(unique(.x)))))) %>%
    mutate(
      across(`Compound Concentration (nM)`, fct_relabel, paste0, " nM")
    ) %>%
    filter(lspci_id %in% measured_in_all) %>%
    group_by(`Compound Concentration (nM)`) %>%
    filter(n() > 10) %>%
    summarize(
      mat = cur_data() %>%
        pivot_wider(names_from = `DiscoveRx Gene Symbol`, values_from = `Percent Control`) %>%
        column_to_rownames("lspci_id") %>%
        as.matrix() %>%
        list(),
      .groups = "drop"
    ) %>% {
      set_names(.$mat, .$`Compound Concentration (nM)`)
    }
  # all_doses_mats
  heatmaps <- map(all_doses_mats, plot_mat, cr = co_cluster_mats(all_doses_mats), cc = co_cluster_mats(map(all_doses_mats, t)))
  # patchwork::wrap_plots(map(heatmaps, "gtable"))
  # egg::ggarrange(
  #   plots = map(heatmaps, "gtable"),
  #   labels = names(heatmaps)
  # )
  cowplot::plot_grid(
    plotlist = map(heatmaps, "gtable"),
    labels = names(heatmaps)
  )
  # do.call(gridExtra::grid.arrange, heatmaps)
  # heatmaps
}
```


```{r}
heatmaps <- generate_mat(
  kinase_stats %>%
    filter(family == "HIP") %>%
    pull(`DiscoveRx Gene Symbol`)
)

dir.create("plots/kinase_families_heatmaps", showWarnings = FALSE)
walk(
  kinase_stats$family %>%
    unique(),
    ~ggsave(
      paste0("plots/kinase_families_heatmaps/", .x, ".pdf"),
      generate_mat(
        kinase_stats %>%
          filter(family == .x) %>%
          pull(`DiscoveRx Gene Symbol`)
      )
    )
)
```
