
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(synExtra)
library(here)
library(fst)
library(data.table)
library(powerjoin)

theme_set(theme_minimal())

synapser::synLogin()
syn <- synDownloader("~/data", .cache = TRUE)
```

Use TAS values based on ChEMBL v27 because that didn't include TAS
values derived from our own Kinomescan data

```{r}
tas_table <- syn("syn25173510") %>%
    read_fst(as.data.table = TRUE)
lsp_target_dictionary <- syn("syn25173506") %>%
    read_fst(as.data.table = TRUE)
single_dose_datasets <- syn("syn26486828") %>%
    read_csv()
pseudo_kds <- syn("syn51080578") %>%
    read_csv()
compound_stats <- syn("syn52624367") %>%
    read_csv()
```

```{r}
tas_reduced <- tas_table %>%
    semi_join(
        lsp_target_dictionary %>%
            filter(organism == "Homo sapiens"),
        by = "lspci_target_id"
    ) %>%
    semi_join(
        single_dose_datasets,
        by = c("lspci_id", "symbol" = "hgnc_symbol")
    )

tas_reduced %>%
    group_by(lspci_id, symbol) %>%
    arrange(lspci_id, symbol) %>%
    filter(n() > 1) %>%
    print(n = Inf)
```

Produce clustered heatmap of TAS values that were available for the
OKL before we did Kinomescan.

```{r}
library(seriation)
library(impute)

tas_colors <- c(`1` = "#b2182b", `2` = "#ef8a62", `3` = "#fddbc7", `10` = "#2166ac", `no data` = "#ffffff")

cluster_df <- function(df, row_var, col_var, value_var) {
    mat <- df %>%
        distinct({{row_var}}, {{col_var}}, {{value_var}}) %>%
        pivot_wider(names_from = {{col_var}}, values_from = {{value_var}}) %>%
        column_to_rownames(rlang::as_name(rlang::enquo(row_var)))
    mat_imp <- impute.knn(
        t(mat), colmax = 0.9999
    ) %>%
        chuck("data") %>%
        t()
    dist_rows <- dist(mat_imp, method = "euclidian")
    dist_cols <- dist(t(mat_imp), method = "euclidian")
    clust_rows <- hclust(dist_rows, method = "average") %>%
        reorder(dist_rows, method = "olo")
    clust_cols <- hclust(dist_cols, method = "average") %>%
        reorder(dist_cols, method = "olo")
    df %>%
        mutate(
            "{{row_var}}" := factor({{row_var}}, levels = clust_rows$labels[clust_rows$order]),
            "{{col_var}}" := factor({{col_var}}, levels = clust_cols$labels[clust_cols$order])
        )
}

tas_heatmap <- function(df, row_var, col_var, value_var) {
    df %>%
        ggplot(aes({{col_var}}, {{row_var}})) +
        geom_raster(aes(fill = {{value_var}})) +
        scale_fill_manual(values = tas_colors, na.value = "white") +
        theme_minimal()
}

tas_reduced_clustered <- cluster_df(tas_reduced, lspci_id, symbol, tas) %>%
    mutate(tas = fct_inseq(as.character(tas)) %>% fct_na_value_to_level("no data"))
p <- tas_heatmap(tas_reduced_clustered, lspci_id, symbol, tas) +
    labs(x = "Kinase", y = "Compound", fill = "TAS") +
    coord_equal() +
    theme_bw() +
    theme(
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank()
    )

ggsave(
    "plots/tas_heatmap.pdf",
    p, width = 7, height = 3.5
)
```

Do not need to filter by discordant here. Was already done before calculating
pseudo Kd values.

```{r}
pseudo_kds_tas <- pseudo_kds %>%
    filter(
        dataset == "original_repeat_replaced",
        !exclude_target
    ) %>%
    semi_join(
        compound_stats %>%
            filter(standard_doses_measured),
        by = c("dataset", "lspci_id")
    ) %>%
    mutate(
        tas_pseudo_kd = cut(
            pseudo_kd,
            breaks = c(-Inf, 100, 999, 9999, Inf),
            labels = c("1", "2", "3", "10")
        )
    ) %>%
    group_by(lspci_id, hgnc_symbol) %>%
    arrange(tas_pseudo_kd) %>%
    slice_head(n = 1) %>%
    ungroup()

pseudo_kds_tas_clustered <- pseudo_kds_tas %>%
    mutate(
        hgnc_symbol = factor(hgnc_symbol, levels = levels(tas_reduced_clustered$symbol)),
        lspci_id = factor(lspci_id, levels = levels(tas_reduced_clustered$lspci_id))
    )

p <- tas_heatmap(pseudo_kds_tas_clustered, lspci_id, hgnc_symbol, tas_pseudo_kd) +
    labs(x = "Kinase", y = "Compound", fill = "TAS") +
    coord_equal() +
    theme_bw() +
    theme(
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank()
    )

pseudo_kds_tas_clustered_2 <- pseudo_kds_tas %>%
    mutate(across(tas_pseudo_kd, \(x) as.numeric(as.character(x)))) %>%
    cluster_df(lspci_id, hgnc_symbol, tas_pseudo_kd) %>%
    mutate(across(tas_pseudo_kd, \(x) factor(as.character(x), levels = c("1", "2", "3", "10"))))

pseudo_kds_tas_clustered_2 %>%
    complete(lspci_id, hgnc_symbol) %>%
    group_by(hgnc_symbol) %>%
    summarize(
        n = sum(is.na(tas_pseudo_kd)),
        .groups = "drop"
    ) %>%
    arrange(desc(n))

tas_reduced_clustered_2 <- tas_reduced_clustered %>%
    filter(
        lspci_id %in% pseudo_kds_tas_clustered_2$lspci_id,
        symbol %in% pseudo_kds_tas_clustered_2$hgnc_symbol
    ) %>%
    mutate(
        lspci_id = factor(lspci_id, levels = levels(pseudo_kds_tas_clustered_2$lspci_id)),
        symbol = factor(symbol, levels = levels(pseudo_kds_tas_clustered_2$hgnc_symbol))
    ) %>%
    group_by(lspci_id, symbol) %>%
    arrange(tas) %>%
    slice_head(n = 1) %>%
    ungroup()

tas_clustered_both_2 <- bind_rows(
    OKL = pseudo_kds_tas_clustered_2 %>%
        mutate(tas = tas_pseudo_kd),
    `Pre-OKL` = tas_reduced_clustered_2 %>%
        mutate(hgnc_symbol = symbol),
    .id = "dataset"
) %>%
    complete(
        dataset, lspci_id, hgnc_symbol, fill = list(tas = "no data")
    )

p <- tas_heatmap(tas_clustered_both_2, lspci_id, hgnc_symbol, tas) +
    labs(x = "Kinase", y = "Compound", fill = "TAS") +
    coord_equal() +
    theme_bw() +
    facet_wrap(~dataset) +
    guides(fill = guide_legend(override.aes = list(color = "black"))) +
    theme(
        strip.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        legend.key = element_rect(color = "black")
    )

ggsave(
    "plots/tas_heatmap_pre_okl_vs_okl_alternate_colors.pdf",
    p, width = 10, height = 4
)

tas_heatmaps <- tas_clustered_both_2 %>%
    group_nest(dataset) %>%
    mutate(
        p = map(
            data,
            \(x) tas_heatmap(x, lspci_id, hgnc_symbol, tas)+
                labs(x = "Kinase", y = "Compound", fill = "TAS") +
                coord_equal() +
                theme_bw() +
                guides(fill = guide_legend(override.aes = list(color = "black"))) +
                theme(
                    strip.background = element_blank(),
                    panel.grid.major = element_blank(),
                    panel.grid.minor = element_blank(),
                    axis.text.x = element_blank(),
                    axis.text.y = element_blank(),
                    axis.ticks = element_blank(),
                    legend.key = element_rect(color = "black")
                )
        )
    )

pwalk(
    ps,
    \(dataset, p, ...) {
        ggsave(
            paste0("plots/tas_heatmap_pre_okl_vs_okl_", dataset, ".pdf"),
            p, width = 5, height = 4
        )
    }
)
```

```{r}
margin_plot <- \(x, m) {
    m_sym <- sym(m)
    x %>%
        mutate(across(tas, fct_rev)) %>%
        group_by(!!m_sym, tas) %>%
        summarize(n = n(), .groups = "drop") %>%
        ggplot(aes(!!m_sym, n, fill = tas)) +
            geom_col() +
            scale_fill_manual(values = tas_colors, guide = "none") +
            theme(
                strip.background = element_blank(),
                panel.grid.major = element_blank(),
                panel.grid.minor = element_blank(),
                # axis.title = element_blank(),
                axis.ticks = element_blank()
            )

}

tas_heatmaps_margins <- tas_heatmaps %>%
    mutate(
        p_col = map(
            data,
            \(x) margin_plot(x, "hgnc_symbol") +
                theme(axis.text.x = element_blank())
        ),
        p_row = map(
            data,
            \(x) margin_plot(x, "lspci_id") +
                coord_flip() +
                scale_y_reverse() +
                theme(axis.text.y = element_blank())
        ),
        p_combined = pmap(
            list(p, p_col, p_row),
            \(p, p_col, p_row) {
                # cowplot::plot_grid(
                #     NULL, p_col,
                #     p_row, p,
                #     nrow = 2, ncol = 2,
                #     rel_heights = c(.2, 1),
                #     rel_widths = c(.4, 1),
                #     align = "hv", axis = "lrtb"
                # )
                cowplot::plot_grid(
                    NULL, p_col + theme(plot.margin = unit(c(0,0,0,0), "pt")) + labs(x = NULL, y = "Frequency"),
                    p_row + theme(plot.margin = unit(c(0,0,0,0), "pt")) + labs(x = "Compound", y = "Frequency"), p + theme(plot.margin = unit(c(0,0,0,0), "pt")) + coord_cartesian(clip = "off") + labs(y = NULL),
                    nrow = 2, ncol = 2,
                    rel_widths = c(.2, 1),
                    rel_heights = c(.2, 1),
                    align = "hv", axis = "lrtb"
                )
            }
        ),
        p_combined2 = pmap(
            list(p, p_col, p_row),
            \(p, p_col, p_row) {
                patchwork::wrap_plots(
                    p_col + theme(plot.margin = unit(c(0,0,0,0), "pt")) + labs(x = NULL, y = NULL),
                    p_row + theme(plot.margin = unit(c(0,0,0,0), "pt")) + labs(x = "Compound", y = "Frequency"), p + theme(plot.margin = unit(c(0,0,0,0), "pt")) + coord_cartesian(clip = "off") + labs(y = NULL),
                    design = "#A\nBC",
                    byrow = TRUE,
                    nrow = 2, ncol = 2,
                    widths = c(.2, 1),
                    heights = c(.2, 1)
                )
            }
        )
    )

pwalk(
    tas_heatmaps_margins,
    \(dataset, p_combined, p_combined2, ...) {
        ggsave(
            paste0("plots/tas_heatmap_pre_okl_vs_okl_", dataset, "_margins.pdf"),
            p_combined2, width = 10, height = 10
        )
    }
)
```


Remaining vertical stripes of NAs are due to kinases like PRKD2, PKN1, and RAF1
which have a large number of measurements at 12.5 and 1000 nM that are below
50% but above 35% remaining and are thus not picked up as discordant. For the
purposes of pseudo Kd they are discordant though (first measurement <50% and
then going up again) so the pseudo Kd is NA.

```{r}
library(ggalluvial)

tas_clustered_both_alluvial <- tas_clustered_both_2 %>%
    mutate(
        dataset = factor(dataset, levels = c("Pre-OKL", "OKL"))
    ) %>%
    select(dataset, lspci_id, hgnc_symbol, tas) %>%
    pivot_wider(names_from = dataset, values_from = tas) %>%
    count(`OKL`, `Pre-OKL`, sort = TRUE, name = "freq") %>%
    mutate(
        added = factor(
            if_else(`OKL` != "no data" & `Pre-OKL` == "no data", "added", "existing"),
            levels = c("added", "existing")
        )
    )

tas_clustered_both_alluvial_long <- tas_clustered_both_alluvial %>%
    to_lodes_form(
        axes = 1:2, key = "Dataset", value = "TAS"
    ) %>%
    mutate(
        across(Dataset, fct_rev)
    )

p <- tas_clustered_both_alluvial_long %>%
    mutate(
        Dataset = fct_recode(Dataset, `Original\nTAS` = "Pre-OKL", `Final\nTAS` = "OKL")
    ) %>%
    filter(freq > 200) %>%
    ggplot(
        aes(x = Dataset, y = freq, stratum = TAS, alluvium = alluvium)
    ) +
    scale_x_discrete(expand = c(0.05, 0.05)) +
    geom_alluvium(aes(fill = added), width = 1/10) +
    labs(fill = NULL) +
    scale_fill_manual(
        values = c(existing = "#a2a2a2", added = "#c81919"),
        guide = "none"
    ) +
    ggnewscale::new_scale_fill() +
    scale_fill_manual(values = tas_colors, guide = "none") +
    geom_stratum(aes(fill = TAS), width = 1/6) +
    # geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
    geom_label(stat = "stratum", aes(label = after_stat(stratum)), label.size = 0) +
    theme_minimal() +
    labs(y = "Data points", fill = NA) +
    theme(
        panel.grid.major.x = element_blank()
    ) +
    coord_cartesian(clip = "off")

ggsave(
    "plots/tas_alluvial.pdf",
    p, width = 3, height = 5
)
```


```{r}
library(cvms)

p <- plot_confusion_matrix(
    tas_clustered_both_alluvial,
    target_col = "Pre-OKL",
    prediction_col = "OKL",
    counts_col = "freq"
) +
    labs(x = "Pre-OKL", y = "OKL")

ggsave(
    "plots/tas_confusion_matrix.pdf",
    p, width = 6, height = 6
)
```

How many kinases were assayed for each compound before we did Kinomescan?

```{r}
p <- tas_reduced %>%
    group_by(lspci_id) %>%
    summarize(n_kinases = n_distinct(symbol), ..groups = "drop") %>%
    ggplot(aes(n_kinases)) +
    geom_histogram() +
    labs(x = "Number of kinases assayed", y = "Number of compounds", title = "Initital TAS") +
    theme(plot.title = element_text(hjust = 0.5))

```

