---
title: "OKL single dose data wrangling"
author: "Clemens Hug"
date: "12/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(synExtra)
library(here)
library(fst)
library(data.table)
library(powerjoin)

theme_set(theme_minimal())

synapser::synLogin()
syn <- synDownloader("~/data", .cache = TRUE)
```

## Load input

Trying to find previous Kinomescan runs of OKL compounds.

```{r load}
kinomescan_plan <- syn("syn52433757") %>%
  readxl::read_excel(sheet = "plate_map_metadata_single_dose_") %>%
  dplyr::rename(library = ...1) %>%
  fill(library, .direction = "down")

hmsl_id_cols <- c(
  "HMSLID",
  "HMS ID",
  "Compound Name",
  "Small Molecule HMS LINCS ID"
)

all_kinomescan <- synGlob("syn18502717", "*.csv") %>%
  {.[setdiff(names(.), c("eurofins_kinase_info.csv", "compound_id.csv", "DER009-01-p-00001_SScores Report.csv"))]} %>%
  map(syn) %>%
  map(read_csv)

all_kinomescan_ids <- all_kinomescan %>%
  map(
    \(df) {
      id_col <- hmsl_id_cols[hmsl_id_cols %in% names(df)][1]
      id_col_sym <- sym(id_col)
      df %>%
        mutate(
          hmsl_id_clean = str_replace_all(!!id_col_sym, fixed("-"), "") %>%
          {if_else(!str_starts(., "HMSL"), paste0("HMSL", .), .)}
        ) %>%
        distinct(!!id_col_sym, hmsl_id_clean)
    }
  )

all_kinomescan_ids_overlap <- all_kinomescan_ids %>%
  imap(
    \(df, file) {
      df %>%
        filter(hmsl_id_clean %in% kinomescan_plan$hmsl_id) %>%
        inner_join(all_kinomescan[[file]])
    }
  ) %>%
  keep(\(df) nrow(df) > 0)

all_kinomescan_ids_overlap_long <- all_kinomescan_ids_overlap %>%
  bind_rows(.id = "file")

all_kinomescan_ids_overlap_long %>%
  distinct(hmsl_id_clean, `DiscoveRx Gene Symbol`, `Percent Control`, `Compound Concentration (nM)`)

all_kinomescan_ids_overlap[["SCANmax_results_2.csv"]][["Compound Name"]] %>%
  unique() %>%
  paste(collapse = ", ")
```

"$SCANmax_results_2.csv" (syn18502769) and "DER009-01-p-00001_Data Report.csv"
contain exactly the same results. They contain data for 10 of the pre-OKL
compounds. The remaining 6 are in syn18502768, but are referred to only by their
name.

Updating syn18502768 to also include HMSL id using mapping provided by Caitlin.

```{r}
scanmax1_mapping <- syn("syn18502757") %>%
  read_csv()
scanmax1_raw <- syn("syn18502768") %>%
  read_csv()

scanmax1_mapped <- scanmax1_raw %>%
  power_inner_join(
    scanmax1_mapping %>%
      select(`Compound Name` = `Name`, HMSLID = `HMS ID`),
    by = "Compound Name",
    check = check_specs(
      unmatched_keys_left = "warn",
      unmatched_keys_right = "warn",
      duplicate_keys_right = "warn"
    )
  ) %>%
  select(-`Gray ID`)

fwrite(
  scanmax1_mapped,
  here("data", "SCANmax_results_1.csv")
)

synStoreMany(
  here("data", "SCANmax_results_1.csv"),
  "syn18502717",
  forceVersion = FALSE
)
```

```{r}
inputs <- c(
  compound_mapping = "syn26260389",
  exclusions = "syn26401280",
  pre_okl1 = "syn18502768",
  pre_okl2 = "syn18502769",
  okl1_1 = "syn26388880",
  okl1_2 = "syn25998856",
  okl2 = "syn26053108",
  okl2_repeat1 = "syn52223890",
  okl2_repeat2 = "syn52223891",
  kinase_info = "syn51286743",
  compound_dictionary = "syn26260332",
  kinomescan_plan = "syn52433757"
)

input_files <- map(
  inputs, syn
)

compound_dictionary <- input_files[["compound_dictionary"]] %>%
  read_fst(as.data.table = TRUE)

compound_names <- compound_dictionary %>%
  select(lspci_id, name = pref_name) %>%
  drop_na() %>%
  # lazy_dt() %>%
  group_by(lspci_id) %>%
  slice_head(n = 1) %>%
  ungroup()
  # as_tibble()

compound_mapping <- input_files[["compound_mapping"]] %>%
  read_fst(as.data.table = TRUE)

exclusions <- input_files[["exclusions"]] %>%
  read_csv()

qpcr_misses <- syn("syn33531640") %>%
  readxl::read_excel() %>%
  left_join(
    compound_mapping[, .(lspci_id, external_id)] %>%
      unique(),
    by = c("Compound Name" = "external_id")
  )

okl_single_dose_raw <- input_files %>%
  {.[str_starts(names(.), "okl")]} %>%
  map(syn) %>%
  map(fread) %>%
  rbindlist(idcol = "experiment")

pre_okl_single_dose_raw <- input_files %>%
  {.[str_starts(names(.), "pre_okl")]} %>%
  map(syn) %>%
  map(fread) %>%
  rbindlist(idcol = "experiment", fill = TRUE)

kinase_info <- input_files[["kinase_info"]] %>%
  read_csv()
```

Map HMSL IDs to lspci IDs

```{r}
hmsl_lspci_id_map <- tibble(
  hmsl_id = c(
    okl_single_dose_raw$`Compound Name`,
    pre_okl_single_dose_raw$HMSLID,
    "HMSL11996" # HMSL12274 is deprecated in favor of HMSL11996
  ) %>%
    unique()
) %>%
  power_left_join(
    compound_mapping[, .(lspci_id, hmsl_id = external_id)],
    by = "hmsl_id",
    check = check_specs(
      unmatched_keys_left = "warn",
    )
  )
```

## Wrangle

We ran KinomeScan assays against the entire Kinome for all compounds in the
Optimal Kinase Library. Experiments were performed in multiple rounds. Some of
the OKL compounds were already run in 2019 before OKL was conceived and thus
were not run again in the main OKL experiments.

Some compounds in these runs were also not part of the OKL. These were
JAK inhibitors or other compounds ran at the same time. Adding a `library` column
to distinguish compounds from these different sources.

Some data that was generated in the main round of OKL experiments turned out to
be bad and was repeated.

```{r wrangle}
compound_library_map <- kinomescan_plan %>%
  drop_na(hmsl_id) %>%
  distinct(library, hmsl_id) %>%
  mutate(
    library = if_else(str_detect(library, fixed("OKL")), "OKL", "JAKi")
  ) %>%
  bind_rows(
    tibble(
      hmsl_id = setdiff(hmsl_lspci_id_map$hmsl_id, .$hmsl_id),
      library = "other"
    )
  )

single_dose_lspci_id <- bind_rows(
  pre_okl_single_dose_raw %>%
    select(-`Compound Name`) %>%
    dplyr::rename(hmsl_id = HMSLID),
  okl_single_dose_raw %>%
    dplyr::rename(hmsl_id = `Compound Name`)
) %>%
  # Fix deprecated HMSLID
  mutate(hmsl_id = recode(hmsl_id, HMSL12274 = "HMSL11996")) %>%
  power_left_join(
    hmsl_lspci_id_map,
    by = "hmsl_id",
    check = check_specs(
      unmatched_keys_left = "warn"
    )
  ) %>%
  power_left_join(
    compound_names,
    by = "lspci_id",
    check = check_specs(
      unmatched_keys_left = "warn"
    )
  ) %>%
  as_tibble() %>%
  power_left_join(
    compound_library_map,
    by = "hmsl_id",
    check = check_specs(
      unmatched_keys_left = "warn",
      duplicate_keys_right = "warn"
    )
  )
```

Check and visualize how often each compound was measured at each concentration.

```{r}
single_dose_lspci_id %>%
  dplyr::count(experiment, lspci_id, `Compound Concentration (nM)`) %>%
  dplyr::count(experiment, n)
```
     experiment   n  nn
1:       okl1_1 468 368
2:       okl1_2 468 368
3:         okl2 468  48
4: okl2_repeat1   7 736
5: okl2_repeat2 468   2
6:     pre_okl1 468  23
7:     pre_okl2 468  26

For stats, remove experiment where seven compounds were measured only against
subset of kinases.

```{r}
compound_stats <- single_dose_lspci_id %>%
  filter(experiment != "okl2_repeat1") %>%
  group_by(library, lspci_id, name, `Compound Concentration (nM)`) %>%
  summarize(
    n_measured = n() / 468L,
    .groups = "drop"
  )

compound_stats %>%
  dplyr::count(n_measured)

p <- compound_stats %>%
  filter(
    `Compound Concentration (nM)` %in% c(12.5, 100, 1000, 10000),
    library != "other"
  ) %>%
  mutate(
    across(
      c(lspci_id, `Compound Concentration (nM)`),
      \(x) fct_inseq(as.character(x))
    )
  ) %>%
  complete(fill = list(n_measured = 0)) %>%
  ggplot(
    aes(x = `Compound Concentration (nM)`, y = name, fill = n_measured)
  ) +
  geom_tile()

ggsave(
  "plots/compounds_n_measured.pdf", p,
  width = 6, height = 20
)
```

Checking which compounds were measured multiple times at the same concentration

```{r}
single_dose_lspci_id_repeat <- single_dose_lspci_id %>%
  filter(str_detect(experiment, "repeat"))

single_dose_lspci_id_repeat %>%
  group_by(lspci_id, `DiscoveRx Gene Symbol`, `Compound Concentration (nM)`) %>%
  filter(n() > 1)
```

Erlotinib has some duplicate data values because it was measured twice. Once
in the experiment where it was measured against all kinases and then in the
experiment where certain kinases where measured against all compounds.

Averaging measurements.

```{r}
single_dose_lspci_id_no_repeat <- single_dose_lspci_id %>%
  filter(!str_detect(experiment, "repeat"))

single_dose_lspci_id_no_repeat %>%
  group_by(lspci_id, `DiscoveRx Gene Symbol`, `Compound Concentration (nM)`) %>%
  filter(n() > 1)

single_dose_lspci_id_no_repeat %>%
  group_by(lspci_id, `DiscoveRx Gene Symbol`, `Compound Concentration (nM)`) %>%
  filter(n() > 1) %>% ungroup() %>% dplyr::count(experiment, name)
```

Tofacitinib and Torin-1 were measured twice. Averaging measurements.

Before averaging, flagging measurements for exclusion for multiple reasons.

QPCR misses and mutations are flagged for exclusion.

```{r}
single_dose_processed_all <- bind_rows(
  single_dose_lspci_id_no_repeat %>%
    filter(str_detect(experiment, "pre_okl")),
  single_dose_lspci_id_no_repeat %>%
    filter(!str_detect(experiment, "pre_okl")) %>%
    left_join(
      qpcr_misses %>%
        transmute(
          lspci_id, `Ambit Gene Symbol`, `Compound Conc`,
          qpcr_miss = TRUE
        ),
      by = c("lspci_id", "DiscoveRx Gene Symbol" = "Ambit Gene Symbol", "Compound Concentration (nM)" = "Compound Conc")
    ),
  single_dose_lspci_id_repeat
) %>%
  replace_na(list(qpcr_miss = FALSE)) %>%
  mutate(
    experiment_group = if_else(
      str_detect(experiment, "repeat"),
      "repeat",
      "original"
    )
  )

single_dose_processed_deduped <- single_dose_processed_all %>%
  group_by(
    experiment_group, lspci_id, hmsl_id, name, `DiscoveRx Gene Symbol`, `Entrez Gene Symbol`,
    `Compound Concentration (nM)`
  ) %>%
  summarize(
    experiment = paste(experiment, collapse = "+"),
    library = unique(library),
    `Percent Control` = mean(`Percent Control`[if (all(qpcr_miss)) TRUE else !qpcr_miss]),
    qpcr_miss = all(qpcr_miss),
    .groups = "drop"
  )

add_kinase_info <- function(data) {
  data %>%
    mutate(
      exclude_target = `DiscoveRx Gene Symbol` %in% exclusions$`DiscoveRx Gene Symbol`
    ) %>%
    mutate(
      tas = case_when(
        `Compound Concentration (nM)` == 10000 & `Percent Control` >= 50 ~ 10L,
        `Compound Concentration (nM)` == 10000 & `Percent Control` < 0.1 ~ 2L,
        `Compound Concentration (nM)` == 1000 & `Percent Control` >= 90 ~ 10L,
        `Compound Concentration (nM)` == 1000 & `Percent Control` < 1 ~ 2L,
        `Compound Concentration (nM)` == 100 & `Percent Control` >= 75 ~ 10L,
        `Compound Concentration (nM)` == 100 & `Percent Control` < 25 ~ 2L,
        TRUE ~ NA_integer_
      )
    ) %>%
    power_left_join(
      kinase_info %>%
        select(`DiscoveRx Gene Symbol`, entrezgene_id, ensembl_gene_id, hgnc_symbol),
      by = "DiscoveRx Gene Symbol",
      check = check_specs(
        unmatched_keys_left = "warn",
        duplicate_keys_right = "warn"
      )
    ) %>%
    select(
      lspci_id, hmsl_id, library, name,
      `DiscoveRx Gene Symbol`, `Entrez Gene Symbol`,
      entrezgene_id, ensembl_gene_id, hgnc_symbol, exclude_target,
      experiment_group, experiment, qpcr_miss, everything()
    )
}

single_dose_processed_all_final <- single_dose_processed_all %>%
  add_kinase_info()

single_dose_processed_deduped_final <- single_dose_processed_deduped %>%
  add_kinase_info()

fwrite(single_dose_processed_deduped_final, here("data", "okl_single_dose.csv.gz"))
fwrite(single_dose_processed_all_final, here("data", "okl_single_dose_no_agg.csv.gz"))
```

Making two complete datasets:

1. Measurements before the repeat experiment but not including the repeat
  experiments.
2. All measurements, but excluding measurements that were repeated in the
  repeat experiment.

This way it should be easy to compare these two complete datasets in follow-up
analysis.

```{r}
single_dose_datasets <- bind_rows(
  only_original = single_dose_processed_deduped_final %>%
    filter(experiment_group == "original"),
  original_repeat_replaced = single_dose_processed_deduped_final %>%
    filter(experiment_group == "original") %>%
    anti_join(
      single_dose_processed_deduped_final %>%
        filter(experiment_group == "repeat"),
      by = c(
        "lspci_id", "hmsl_id", "DiscoveRx Gene Symbol", "Entrez Gene Symbol",
        "Compound Concentration (nM)"
      )
    ) %>%
    bind_rows(
      single_dose_processed_deduped_final %>%
        filter(experiment_group == "repeat")
    ),
  .id = "dataset"
)

fwrite(single_dose_datasets, here("data", "okl_single_dose_datasets.csv.gz"))
# single_dose_datasets <- fread(here("data", "okl_single_dose_datasets.csv.gz"))
```

Find for which compounds we have all four measurements.

```{r}
compound_stats_datasets <- single_dose_datasets %>%
  group_by(dataset, lspci_id, name, `Compound Concentration (nM)`) %>%
  summarize(
    all_measured = (n() / 468) >= 1,
    .groups = "drop"
  ) %>%
  arrange(dataset, `Compound Concentration (nM)`) %>%
  pivot_wider(
    names_from = `Compound Concentration (nM)`,
    values_from = all_measured,
    values_fill = list(all_measured = FALSE)
  ) %>%
  mutate(
    standard_doses_measured = `12.5` & `100` & `1000` & `10000`
  )

fwrite(
  compound_stats_datasets,
  here("data", "okl_single_dose_datasets_compound_stats.csv.gz")
)
```

## Upload to Synapse

```{r synapse}
# activity <- synapser::Activity(
#   name = "Wrangle OKL KinomeScan data",
#   used = unname(inputs),
#   executed = "https://github.com/labsyspharm/okl-analysis/blob/main/wrangle_data.Rmd"
# )

c(
  here("data", "okl_single_dose.csv.gz"),
  here("data", "okl_single_dose_datasets.csv.gz"),
  here("data", "okl_single_dose_no_agg.csv.gz"),
  here("data", "okl_single_dose_datasets_compound_stats.csv.gz")
) %>%
  synStoreMany(
    parentId = "syn18508401",
    activityName = "Wrangle OKL KinomeScan data",
    used = unname(inputs),
    executed = "https://github.com/labsyspharm/okl-analysis/blob/main/wrangle_data.Rmd",
    forceVersion = FALSE
  )
```
