
```{r}
library(tidyverse)
library(synExtra)
library(fst)
library(data.table)
library(powerjoin)

synapser::synLogin()

syn <- synDownloader("~/data", .cache = TRUE)
```

```{r}
compound_dict <- syn("syn26260332") %>%
  read_fst()

compound_names <- syn("syn26260347") %>%
  read_fst(as.data.table = TRUE)

canonical_names <- compound_names[
  !is.na(name), .(lspci_id, name)
][
  , head(.SD, n = 1), by = .(lspci_id)
]

compound_mapping <- syn("syn26260389") %>%
  read_fst(as.data.table = TRUE)
```

Extracting IC50s from Klaeger data:

"Kinobeads Drugmatrix detailed"
	Column	Description
	Gene name	Gene name or Uniprot Accession of the target.
	Abemaciclib	This and all subsequent columns contain the apparent Kds [nM] of the respective inhibitors.

	Cell content:
	123	This protein-drug interaction was classified as "high-confidence" target (see Supplementary Information for details).
	(123)	This protein-drug interaction was classified as "low-confidence" target (see Supplementary Information for details).
	n.i.	Protein was identified and quantified but the dose-response curve did not suggest a dose-dependent effect and is thus annotated as "not inhibited".
	n.d.	Protein was either not identified or not quantified in this profiling, thus no statement about its inhibition can be made ("not defined").

A protein was considered a high-confidence target if the binding curve showed a sigmoidal shape with a dose-dependent decrease in binding to the Kinobeads.

So confidence here is in the data quality, not in how strongly the compound binds.

```{r}
klaeger_ic50s_raw <- syn("syn51288788") %>%
    readxl::read_excel(sheet = "Kinobeads Drugmatrix detailed")

klaeger_ic50s <- klaeger_ic50s_raw %>%
    pivot_longer(
        -c(Name, Kinase, `Direct binder`),
        names_to = "Compound",
        values_to = "IC50_raw"
    ) %>%
    mutate(
        IC50 = case_when(
            str_detect(IC50_raw, "[0-9]+") ~ as.double(str_extract(IC50_raw, "[0-9]+")),
            IC50_raw == "n.i." ~ 5e4,
            TRUE ~ NA_real_
        ) %>%
            pmin(5e4),
        IC50_relation = case_when(
            IC50 >= 5e4 ~ ">",
            IC50 < 5e4 ~ "=",
            TRUE ~ NA_character_
        ),
        IC50_quality = case_when(
            !str_detect(IC50_raw, "[0-9]+") ~ NA_character_,
            str_detect(IC50_raw, "\\(") ~ "low",
            TRUE ~ "high"
        )
    )

klaeger_single_dose_raw <- syn("syn51288926") %>%
    readxl::read_excel(sheet = "Kinobeads")


setdiff(klaeger_ic50s$Compound, compound_names$name)

setdiff(
  str_replace(klaeger_ic50s$Compound, fixed("-"), ""),
  str_replace(compound_names$name, fixed("-"), "")
)

klaeger_mapping <- klaeger_ic50s %>%
  distinct(Compound) %>%
  mutate(name_join = str_replace(Compound, fixed("-"), "") %>% str_to_lower())

klaeger_mapping %>%
  arrange(name_join) %>%
  group_by(name_join) %>%
  filter(n() > 1)

klaeger_mapping <- klaeger_mapping %>%
  power_left_join(
    compound_names %>%
      drop_na(name) %>%
      mutate(name = str_replace(name, fixed("-"), "") %>% str_to_lower()) %>%
      filter(name %in% klaeger_mapping$name_join) %>%
      distinct() %>%
      group_by(name) %>%
      filter(if (any(priority == "primary")) priority == "primary" else TRUE) %>%
      arrange(lspci_id) %>%
      slice_head(n = 1) %>%
      ungroup() %>%
      distinct(lspci_id, name),
    by = c("name_join" = "name"),
    check = check_specs(
      duplicate_keys_left = "abort",
      duplicate_keys_right = "abort"
    )
  ) %>%
  power_left_join(
    compound_dict %>%
      select(lspci_id, hmsl_id, chembl_id),
    by = c("lspci_id"),
    check = check_specs(
      duplicate_keys_left = "warn",
      duplicate_keys_right = "abort"
    ),
    na_matches = "never"
  )

klaeger_ic50s_matched <- klaeger_ic50s %>%
  power_left_join(
    klaeger_mapping %>%
      select(Compound, lspci_id, hmsl_id, chembl_id),
    by = "Compound",
    check = check_specs(
      duplicate_keys_right = "abort",
      unmatched_keys_right = "abort"
    )
  ) %>%
  select(Compound, lspci_id, hmsl_id, chembl_id, everything())

write_csv(
    klaeger_ic50s_matched,
    file.path("data", "klaeger_ic50s.csv.gz")
)
synStoreMany(
    file.path("data", "klaeger_ic50s.csv.gz"),
    parentId = "syn45321440",
    forceVersion = FALSE
)


klaeger_single <- klaeger %>%
  select(Drug, lspci_id, hmsl_id, chembl_id, `Gene Name`, starts_with("Relative Intensity ")) %>%
  pivot_longer(
    starts_with("Relative Intensity "),
    names_pattern = "Relative Intensity ([0-9]+) nM",
    names_to = "Concentration", values_to = "Relative Intensity"
  ) %>%
  # DMSO
  drop_na(Concentration, `Relative Intensity`) %>%
  mutate(across(Concentration, as.double))

klaeger_single %>%
  filter(Drug == "ACTB-1003", `Gene Name` == "BCR") %>%
  ggplot(aes(Concentration, `Relative Intensity`)) +
    geom_point() +
    geom_line() +
    scale_x_log10()
```

